Assessing Tor's Usability as a Censorship Circumvention Tool

### 2015-05-12 cs261n-paper/paper.tex
Remove extra space before refs.

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot Tor usability study to test the usability of Tor as an anonymity tool,
targeting a specific demographic of users known to use Tor (journalists who have reported in other
countries) and aiming to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, we refer you to Appendix ~\ref{sec:pilot}.

@@

Each session of 30 participants will be placed in the same simulated censorship environment. 
We start off the experiment by informing our participants that they are in an adversarial setting, 
where some websites are blocked while other websites are not. We will instruct them to visit a censored
website and a non-censored website to illustrate that their devices and the network is fully functional and
operating as expected, but that they are in a censorship environment where some websites will fail to load. 
We will take a moment to explain what Tor is, how it is generally used in this case, and ask them to complete 
the aforementioned set of tasks. To complete all the tasks, the participants will ultimately need to get the correct
configuration settings for the country they are simulated to be in.  Figure ~\ref{dialog} shows one of the 
Tor configuration dialog screens that a user will need to navigate.

@@

The survey consists of 17 questions total, with a basic breakdown as follows. 
The complete list of survey questions can be found in Appendix ~\ref{sec:survey}.

@@

\subsection{Usability for Security Software}
\indent \indent 
We evaluate the usability of the configuration window's interface. Generally, usability 
refers to the learnability, efficiency, memorability, error rate/recoverability, and satisfaction
during use. To achieve these goals, the HCI community has established heuristics for user
interface design: visibility of the system status, using familiar
language, preventing errors, having consistent visuals, flexibility and efficiency of use, 
error recovery, and leveraging recognition over memorization ~\cite{nielsen199510}. Note
that these are not specific usability guidelines, but heuristics.

Usability has sligtly different meanings depending on the context. Some emphasize
learnability, flexibility, and ease of user. In a security context, the priorities are whatever is 
required for the security to be used effectively. As stated by Whitten \etal, a security software is 
usable if people who are expected to use it are reliably made aware of the security tasks to perform, 
are able to figure out how to successfully perform those tasks, do not make dangerous errors, and
are sufficiently comfortable with the software interface ~\cite{whitten1999johnny}. Since usable security 
has user interface design goals than general consumer software, it likewise requires usability evaluation
methods which are suitable to test these properties. Our goals are to test that the Tor Browser, 
communicates an accurate conceptual model of the task to the users as quickly as possible, provide guidance to 
users at the right time, and use warnings and other interactive tools to prevent any mistakes.

\subsection{Usability and Tor}
\indent \indent  Work by Clark \etal to explore ways to simplify the Tor user interface
has led to great usability improvements ~\cite{clark2007usability}.  Clark analyzed the challenges 
of using multiple Tor-related tools including TorPark, Vidalia, FoxyProxy, and TorButton to find that
none are fully satisfactory from a usability perspective. Researchers provide guidelines on 
how to incorporate the best aspects of each tool, while providing a set of guidelines drawn from
usable security and computer human interaction. Key takeaways from this work includes 
finding out that exposure to how Tor works may be difficult for users to understand, and that 
familiar, non-technical language is critical for communication. We plan to draw from Clark's
guidelines as well as other works in usable security and computer human interaction to suggest
any improvements to the Tor Browser.

This will be the first experiment to evaluate Tor's usability as a censorship circumvention tool.
Previous work on evaluating the usability of Tor as an anonymity system have been fruitful. 
Norcie \etal identified ``stop points,'' or barriers to installing and using Tor to help streamline
the interface. Key stop points included the ability to discriminate between anonymized and non-
anonymized browser windows while using tor, unclear download and install processes, and 
confusion as to why the Tor behaves in peculiar ways ~\cite{norcie2012eliminating}. A follow up
study by Norcie \etal verifies that the changes recommended to the Tor Project were indeed 
effective at reducing stopping points mentioned in the previous work. The authors also present
design heuristics for designing usable anonymous systems, such as easy installation, communicating
tradeoffs made to users, and informing why, not how, precautions are taken for security guarantees ~\cite{norcie2014johnny}.

@@

To first ensure that our users knew what a cognitive walkthrough is, we explain what a cognitive
walkthrough and give a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then ask our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough along the process of completing the task. We give our 
participants feedback on how to properly conduct a cognitive walkthrough ~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.

>>>>

\section{A Pilot Tor Usability Study} 
\indent \indent We ran a pilot Tor usability study to test the usability of Tor as an anonymity tool,
targeting a specific demographic of users known to use Tor (journalists who have reported in other
countries) and aiming to identify how well they could understand Tor's non-typical behaviors and 
navigate advanced settings. For the full details of this pilot experiment, we refer you to Appendix~\ref{sec:pilot}.

@@

Each session of 30 participants will be placed in the same simulated censorship environment. 
We start off the experiment by informing our participants that they are in an adversarial setting, 
where some websites are blocked while other websites are not. We will instruct them to visit a censored
website and a non-censored website to illustrate that their devices and the network is fully functional and
operating as expected, but that they are in a censorship environment where some websites will fail to load. 
We will take a moment to explain what Tor is, how it is generally used in this case, and ask them to complete 
the aforementioned set of tasks. To complete all the tasks, the participants will ultimately need to get the correct
configuration settings for the country they are simulated to be in.  Figure~\ref{dialog} shows one of the 
Tor configuration dialog screens that a user will need to navigate.

@@

The survey consists of 17 questions total, with a basic breakdown as follows. 
The complete list of survey questions can be found in Appendix~\ref{sec:survey}.

@@

\subsection{Usability for Security Software}
\indent \indent 
We evaluate the usability of the configuration window's interface. Generally, usability 
refers to the learnability, efficiency, memorability, error rate/recoverability, and satisfaction
during use. To achieve these goals, the HCI community has established heuristics for user
interface design: visibility of the system status, using familiar
language, preventing errors, having consistent visuals, flexibility and efficiency of use, 
error recovery, and leveraging recognition over memorization~\cite{nielsen199510}. Note
that these are not specific usability guidelines, but heuristics.

Usability has sligtly different meanings depending on the context. Some emphasize
learnability, flexibility, and ease of user. In a security context, the priorities are whatever is 
required for the security to be used effectively. As stated by Whitten \etal, a security software is 
usable if people who are expected to use it are reliably made aware of the security tasks to perform, 
are able to figure out how to successfully perform those tasks, do not make dangerous errors, and
are sufficiently comfortable with the software interface~\cite{whitten1999johnny}. Since usable security 
has user interface design goals than general consumer software, it likewise requires usability evaluation
methods which are suitable to test these properties. Our goals are to test that the Tor Browser, 
communicates an accurate conceptual model of the task to the users as quickly as possible, provide guidance to 
users at the right time, and use warnings and other interactive tools to prevent any mistakes.

\subsection{Usability and Tor}
\indent \indent  Work by Clark \etal to explore ways to simplify the Tor user interface
has led to great usability improvements~\cite{clark2007usability}.  Clark analyzed the challenges 
of using multiple Tor-related tools including TorPark, Vidalia, FoxyProxy, and TorButton to find that
none are fully satisfactory from a usability perspective. Researchers provide guidelines on 
how to incorporate the best aspects of each tool, while providing a set of guidelines drawn from
usable security and computer human interaction. Key takeaways from this work includes 
finding out that exposure to how Tor works may be difficult for users to understand, and that 
familiar, non-technical language is critical for communication. We plan to draw from Clark's
guidelines as well as other works in usable security and computer human interaction to suggest
any improvements to the Tor Browser.

This will be the first experiment to evaluate Tor's usability as a censorship circumvention tool.
Previous work on evaluating the usability of Tor as an anonymity system have been fruitful. 
Norcie \etal identified ``stop points,'' or barriers to installing and using Tor to help streamline
the interface. Key stop points included the ability to discriminate between anonymized and non-
anonymized browser windows while using tor, unclear download and install processes, and 
confusion as to why the Tor behaves in peculiar ways~\cite{norcie2012eliminating}. A follow up
study by Norcie \etal verifies that the changes recommended to the Tor Project were indeed 
effective at reducing stopping points mentioned in the previous work. The authors also present
design heuristics for designing usable anonymous systems, such as easy installation, communicating
tradeoffs made to users, and informing why, not how, precautions are taken for security guarantees~\cite{norcie2014johnny}.

@@

To first ensure that our users knew what a cognitive walkthrough is, we explain what a cognitive
walkthrough and give a demonstration of a cognitive walkthrough on a non-related dummy task 
(drawing a red hexagon in the computer graphics program Paint). We then ask our users to perform
an unrelated Internet-based task (finding the population of Zimbabwe) using any browser of their choice
while performing a cognitive walkthrough along the process of completing the task. We give our 
participants feedback on how to properly conduct a cognitive walkthrough~\cite{wharton1994cognitive}.
This process was also used to get users comfortable with talking our loud with the researcher and gauge 
how comfortable users were with a computer.
